\magnification=\magstep1
\baselineskip = 20pt
\tolerance=10000
\raggedright
\raggedbottom
\nopagenumbers

\font\big=cmr12 
\font\smallrm=cmr8
\font\smcaps=cmcsc10

\pageno=1
\input bffonts.tex

\def\BIC{\hbox{\rm BIC}}
\def\Var{\hbox{\rm Var}}

\pageno=1
\centerline{\bf Diagnostic Checking Periodic Autoregression Models}
\centerline{\bf  With Application}
\bigskip
\centerline{A.I. McLeod}
\bigskip
\centerline{{\it The Journal of Time Series Analysis\/}, Vol. 15, No. 2, p.225}

\bigskip

\centerline{\big ADDENDUM}
\bigskip

I have written Splus functions for the methodology discussed in
my paper.
S style documentation for each function is provided.
Splus version 3.2 is required.
To obtain, e-mail the following message:
{\it send pear from S} to statlib@temper.stat.cmu.edu.

I did not precisely define the sample periodic
autocovariance function in the case where the data lengths
for periods are unequal. For example, in the data used
for the Fraser River from March 1913 to December 1991, there
are 78 observations in January and February and 79 in each of
the other months. 
This omission is rectified by noting that in
eqn. (2.2), the $n$ refers to the ceiling of $N/s$ and
the summation is over all data values in the sample.

Note the following typographical corrections:

\item{(i)} p.222, l.22: from March 1912 to December 1990. Figure
1 abscissa labels should be changed accordingly.
\bigskip

\item{(ii)} p.223, l.12: parallel lines at $\pm 1.96/\surd{n}$
\bigskip

\item{(iii)} 
$$\BIC_m = n \ln \hat \sigma_m^2 + \ln(n) p_m.\eqno{(2.4)}$$

\item{(iv)}
$$I_m = {1\over \sigma_m^2} (\gamma_{i-j,m-j}).\eqno{(3.3)}$$


\item{(v)} 
$$\tilde Q_{L,m} = \sum\limits_{\ell=1}^L 
{ \hat r^2_{\ell,m}\over  \Var(r_{\ell,m})  }. \eqno{(4.5)}$$

\item{(vi)}
p.228, l.9
with $(i,j)$ entry $-\psi_{i-j, m-j} \sigma_{m-i} / \sigma_{m}$. 

\item{(vii)} There are several minor numerical errors in 
Tables I and Table III due to using a divisor of 78
instead of 79 for some months.
The the actual fit of the suggested model is slightly improved.
The corrected tables are given in the files table1.tex and table3.tex
which are included with the software available from statlib.


My thanks to Professor A.S. Kozek for pointing out the need to
precisely define the sample periodic autocorrelations in the unequal
lengths case and for bringing the need for some typographical 
corrections to my attention.

\bye
